{\pkg}: A Python Package for Integrating Computational Models Across Languages and Scales

### 2019-01-01 papers/lang2018/ms.tex
Response to reviewer 2 and addition of example in response to reviewer 1

%---------------------------------------------------------------------------------
% PREAMBLE
\iftwocol
	\ifieee
		\documentclass[journal]{IEEEtran}
	\else
		\documentclass[apj]{emulateapj}
	\fi
\else
	\ifieee
		\documentclass[peerreview]{IEEEtran}
	\else
		\documentclass[preprint]{aastex}
	\fi
\fi

\pdfoutput=1
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{listings}
\ifdraft
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\else
		\usepackage{natbib}
		\citestyle{apj}
		\bibliographystyle{apj}
		\bibpunct[,]{(}{)}{;}{a}{}{,}
	\fi
\else
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\fi
\fi

@@

%---------------------------------------------------------------------------------
% CONTROL THINGS
\newcommand{\todo}[1]{{\color{red}{#1}}}
\newcommand{\addref}{{\color{red}(ref. needed)}}
\newcommand{\Msol}{\,M$_{\odot}$}
\newcommand{\TB}{\,Tbytes}
\newcommand{\cis}{{\tt cis\_interface}{}}

@@

\ifieee
\else
	\shorttitle{{\cis}}
\fi

\title{{\cis}: A Python Package for Integrating Computational Models Across Languages and Scales}

@@

%---------------------------------------------------------------------------------
% ABSTRACT (145 words)
\begin{abstract}
Thousands of computational models have been created within the plant biology community in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models. However, the technological barriers introduced by differences in language and data formats has slowed this progress. We present \href{https://github.com/cropsinsilico/cis_interface}{\cis}, a Python package for running integration networks with connections between models across languages and scales. {\cis} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling. {\cis} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
\end{abstract}

@@

Consider the task of integrating a root growth model with a shoot growth model. The root growth model is written in C and can be expressed as
\begin{equation}
R_{t+1} = R_{t} \times r_{r} + R_{t}.
\end{equation}
The inputs to the root growth model are
\begin{itemize}
	\item $R_{t}$: the root mass at time $t$
	\item $r_{r}$: the relative root growth rate
\end{itemize}
and the output of the root growth model is $R_{t+1}$, the root mass at the next time step. 
%
The shoot growth model is written in Python and can be expressed as
\begin{equation}
S_{t+1} = S_{t} \times r_{s} + S_{t} - (R_{t+1} - R_{t}).
\end{equation}
The inputs to the shoot model are 
\begin{itemize}
	\item $S_{t}$: the shoot mass at time $t$
	\item $r_{s}$: the relative shoot growth rate
	\item $R_{t}$: the root mass at time $t$ 
	\item $R_{t+1}$: the root mass at time $t+1$ 
\end{itemize}
and the output of the shoot growth model is $S_{t+1}$, the shoot mass at the next time step.
%
While it is possible to run both models in isolation if all of the appropriate input variables are provided, the value of $R_{t+1}$ calculated by the root growth model could be used as input to the shoot growth model. However, because the two models are written in different languages, the two models cannot be directly integrated.
% without translating one of the models into the language of the other. Although translating one or both of these simple toy models is relatively straight forward, actual plant models are much more complex. In the case of real plant models, model translation is time consuming and results in unnecessary replication of model algorithms.

We present an Open Source Python package, {\cis}, for creating and running integration networks by connecting existing computational models written in different programming language. {\cis} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\cis}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\cis}}. Although developed for the purpose of integrating plant biology models, {\cis} can be applied to any situation requiring coordination between models in the supported languages. \S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\cis} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\cis} package, describes a few use cases, and outlines areas for future development.

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% BACKGROUND
\section{Background}\label{S:background}
%
Computational plant models are usually written with a very specific research question in mind. The resulting programs are highly tuned for one purpose and written specifically for scientists within an isolated group of collaborators. Scientists will generally write programs in the language they are most familiar with and, given that there is no consensus on one programming language as ideal for all scientific applications, the language will often vary between research groups, if not between members within the groups themselves. The same can be said of data formats which can be highly tailored to the target question and may or may not include metadata such as the data types, field names, or units. To integrate two biological models like the root and shoot models from \S\ref{S:intro}, the following questions must be addressed:
%
\begin{enumerate}
	\item Orchestration: How will models be executed?
	\item Communication: How will information be passed from one model to the next across languages?
	\item Translation: How will data output by one model be translated into a format understood by the next model?
\end{enumerate}

These three questions can be addressed through manual integration (e.g. running the root model, converting the output root mass into the units expected by the shoot model, and then running the shoot model). However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than $>3$ models are being integrated. Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models.

@@

{\cis} interface is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}). As a result, models writers only need knowledge of the language in which their model is written.

@@

{\cis} was designed to address the questions from \S\ref{S:background} while being:

@@

{\cis} currently support models written in Python, Matlab, C, and C++ with 
additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).

%---------------------------------------------------------------------------------
% ORCHESTRATION
\subsection{Orchestration}\label{SS:orchestration}
%
{\cis} is executed via a command line interface (CLI), {\tt cisrun} with specification files (\S\ref{SSS:yaml}) as input. Based on the information contained in the specification files, {\cis} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
%At execution, {\cis} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes

@@

Model YAMLs include information about the location of the 
model source code, the language the model is written in, how the model should 
be run, and any input or output variables including their data type and units. 
%
\ifcode
For example, Listings \ref{lst:rootYAML} and \ref{lst:shootYAML} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}. These model YAMLs could be used in any integration without being modified because the model input and output variable will not change.
%
\lstinputlisting[caption={Model YAML for root model},label={lst:rootYAML}]{../../cis_interface/examples/paper/root.yml}
%
\lstinputlisting[caption={Model YAML for shoot model},label={lst:shootYAML}]{../../cis_interface/examples/paper/shoot.yml}
%
\fi
%
Integration networks are specified by declaring the connections 
between models and connections are declared by pairing an output variable from 
one model with the input variable of another model. The {\cis} 
CLI sets up the necessary communication mechanisms to 
then direct data from one model to the next in the specified pattern. Models 
can have as many input and/or output variables as is desired and connections between models are specified by references to the 
input/output variables associated with each model. In addition, input and output variables can also be connected to files.
%
\ifcode
For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:connYAML}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files. The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:rootYAML}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:shootYAML}). The forth connection links the {\tt next\_shoot\_mass} to an output file.
%
\lstinputlisting[caption={Connection YAML for integration of root and shoot models},label={lst:connYAML}]{../../cis_interface/examples/paper/paper.yml}
%
\fi
%
This format allows users the 
flexibility to create complex integration networks and test models in isolation before running the entire integration network.

%---------------------------------------------------------------------------------
% MODEL EXECUTION
\subsubsection{Model Execution}\label{SS:model_execution}
%
{\cis} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly. For example, the root and shoot growth models from \S\ref{S:intro} require 2.42\,s and 4.06\,s respectively to run for 100 time steps in isolation. If these two models were manually integrated in serial via the command line, the integration would require a total of 6.48\,s. However, the same integration requires only 4.52\,s using {\cis}, a speed-up of 1.43. The exact speedup provided by parallel integration using {\cis} depends on how independent the models are and how equally distributed the work is between the models. If the models are fully independent, the speedup is limited by the time required to execute the slowest model. In the example, this would be the shoot growth model which required almost twice the time to execute as the root model. If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be a great.

@@

For the compiled languages (C and C++) there are a few options. The user can 
compile the model themselves, provided they link against the appropriate 
{\cis} header library. Alternatively, users can provide the location of 
the model source code and let {\cis} handle the compilation, including 
linking against the appropriate {\cis} interface library. {\cis} 
also has support for compiling models using Make \citep{Stallman2004} and 
CMake \citep{Martin2006} for 
models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, 
lines are added to the recipe in order to allow linking against {\cis}.

Once the models are running, {\cis} uses threads on the master process to monitor the progress of the model and report back model output (e.g. log messages printed to stdout or stderr) and any status changes. If a model issues any errors, the master process will shut down any model processes that are still running and close any connections, discarding any unprocessed messages. If a model completes without any errors, the master process will cleanup any connections that are no longer required after waiting for all messages to be processed. The master process will only complete once an error is encountered or all model processes have completed.

%---------------------------------------------------------------------------------
% COMMUNICATION
\subsection{Communication}\label{SS:communication}
%
In order for integrations to be possible, models must be able to send and receive information to and from other models written in different languages via some communication mechanism. 
Communication within {\cis} integration networks was designed to be flexible in terms of the 
languages, platforms, and data types available. To accomplish this, {\cis} leverages 
three tools different tools for communication, 
System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), 
which each have their own strengths and weaknesses. The 
particular communication mechanism used by {\cis} is determined by the platform, available libraries, and integration strategy. However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).

%---------------------------------------------------------------------------------
% SYSTEM V IPC
\subsubsection{System V IPC Queues}\label{SSS:ipc}
%
The first communication mechanism used by {\cis} was System V interprocess 
communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
IPC message queues allow messages to be passed between models running on separate 
processes on the same machine. While IPC message queues are simple, fast 
(See \S\ref{SS:results_commtype}), and are built into most Posix operating systems, they do not work in all 
situations. IPC queues are not natively supported by Windows operating systems 
and do not allow communication between remote processes. In addition, IPC queues also 
have relatively low default message size limits on Mac OS X systems (2048 bytes). Once 
the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
sufficient number of messages has been removed from the queue to accommodate the new message. 
For messages larger than the limit, the sending process will block indefinitely. 
This can be handled by splitting large messages into multiple smaller messages 
(See \S\ref{SSS:large}), however, the time required to send a message increases with the number of 
message it must be broken into (See \S\ref{SS:results_commtype}). These limits make sending large messages 
relatively inefficient when compared with other communications mechanisms. 
As a result, IPC queues are used by 
{\cis} as a fallback on Posix systems if none of the other supported communication libraries have 
not been installed.

%---------------------------------------------------------------------------------
% ZEROMQ
\subsubsection{ZeroMQ}\label{SSS:zmq}
%
The preferred communication mechanism used by {\cis} are ZeroMQ sockets 
\citep[ZMQ][]{Akgul2013}. ZMQ 
provides broker-less communication via a number of protocols and patterns with 
bindings in a wide variety of languages that can be installed on Posix and Windows 
operating systems. ZMQ was adopted by {\cis} in order to allow support 
on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
accomplished using IPC queues. In addition, while ZMQ allows 
interprocess communication via IPC, ZMQ also 
supports protocols for distributed communication via an Internet Protocol (IP) 
network. While {\cis} does not currently support using these protocols 
for distributed integration networks, this is a future development plan that has been prepared for.

%---------------------------------------------------------------------------------
% RABBITMQ
\subsubsection{RabbitMQ}\label{SSS:rmq}
%
While broker-less communication like ZMQ is light weight and fast, it is not as fault 
tolerant as brokered messaging systems that confirm message delivery and can resend 
dropped messages. Resilience to dropped messages, while not as necessary for integrations 
running entirely on a local machine, will be more important for integrations running on 
distributed resources with less reliable connections. As a result, {\cis} includes support 
for brokered communication via RabbitMQ \citep[RMQ][]{RMQ} that will be used during future 
development to allow integrations to run on distributed resources or include remote models 
run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
{\cis} does not currently use RMQ for communication unless explicitly specified by the user in their 
integration network YAML.

>>>>

%---------------------------------------------------------------------------------
% PREAMBLE
\iftwocol
	\ifieee
		\documentclass[journal]{IEEEtran}
	\else
		\documentclass[apj]{emulateapj}
	\fi
\else
	\ifieee
		\documentclass[peerreview]{IEEEtran}
	\else
		\documentclass[preprint]{aastex6}
	\fi
\fi

\pdfoutput=1
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[noend]{algpseudocode}
\ifdraft
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\else
		\usepackage{natbib}
		\citestyle{apj}
		\bibliographystyle{apj}
		\bibpunct[,]{(}{)}{;}{a}{}{,}
	\fi
\else
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\fi
\fi

@@

%---------------------------------------------------------------------------------
% CONTROL THINGS
\newcommand{\todo}[1]{{\color{red}{#1}}}
\newcommand{\addref}{{\color{red}(ref. needed)}}
\newcommand{\Msol}{\,M$_{\odot}$}
\newcommand{\TB}{\,Tbytes}
\newcommand{\pkg}{{\tt \todo{cis\_interface}}{}}
\newcommand{\pkgrun}{{\tt \todo{cisrun}}}
\newcommand{\A}{Appendix{ }}

%---------------------------------------------------------------------------------
% SOURCE CODE
\newcommand{\rootsrc}{\lstinputlisting[language=C,label={lst:src:root},caption={root.h: Source code for root model in C.}]{../../cis_interface/examples/root_to_shoot/src/root.h}}
\newcommand{\shootsrc}{\lstinputlisting[language=Python,label={lst:src:shoot},caption={shoot.py: Source code for shoot model in Python.}]{../../cis_interface/examples/root_to_shoot/src/shoot.py}}

%---------------------------------------------------------------------------------
% WRAPPER CODE
\newcommand{\rootwrap}{\lstinputlisting[language=C,label={lst:wrap:root},caption={root\_wrapper.c: Wrapper source code for root model in C.}]{../../cis_interface/examples/root_to_shoot/src/root_wrapper.c}}
\newcommand{\shootwrap}{\lstinputlisting[language=Python,label={lst:wrap:shoot},caption={shoot\_wrapper.py: Wrapper source code for shoot model in Python}]{../../cis_interface/examples/root_to_shoot/src/shoot_wrapper.py}}

%---------------------------------------------------------------------------------
% YAML FILES
\newcommand{\rootyml}{\lstinputlisting[label={lst:yml:root},caption={root.yml: Model YAML specification file for root model.}]{../../cis_interface/examples/root_to_shoot/root.yml}}
\newcommand{\shootyml}{\lstinputlisting[label={lst:yml:shoot},caption={shoot.yml: Model YAML specification file for shoot model.}]{../../cis_interface/examples/root_to_shoot/shoot.yml}}
\newcommand{\rootymlfiles}{\lstinputlisting[label={lst:yml:root_files},caption={root\_files.yml: Connection YAML specification file for running the root model in isolation.}]{../../cis_interface/examples/root_to_shoot/root_files.yml}}
\newcommand{\shootymlfiles}{\lstinputlisting[label={lst:yml:shoot_files},caption={shoot\_files.yml: Model YAML specification file for running the shoot model in isolation.}]{../../cis_interface/examples/root_to_shoot/shoot_files.yml}}
\newcommand{\intyml}{\lstinputlisting[label={lst:yml:int},caption={root\_to\_shoot.yml: Connection YAML specification file for the root/shoot integration.}]{../../cis_interface/examples/root_to_shoot/root_to_shoot.yml}}

%---------------------------------------------------------------------------------
% CODE STYLE
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{1,1,1}
%\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

@@

\ifieee
\else
	\shorttitle{{\pkg}}
\fi

\title{{\pkg}: A Python Package for Integrating Computational Models Across Languages and Scales}

@@

%---------------------------------------------------------------------------------
% ABSTRACT (145 words)
\begin{abstract}
Thousands of computational models have been created within the plant biology community in the past two decades that have the potential to be combined into complex integration networks capable of capturing more complex biological processes than possible with isolated models. However, the technological barriers introduced by differences in language and data formats has slowed this progress. We present \href{https://github.com/cropsinsilico/cis_interface}{\pkg}, a Python package for running integration networks with connections between models across languages and scales. {\pkg} coordinates parallel execution of models in Python, C, C++, and Matlab on Linux, Mac OS, and Windows operating systems, and handles communication in a number of data formats common to computational plant modeling. {\pkg} is designed to be user-friendly and can be accessed at {\tt https://github.com/cropsinsilico/cis\_interface}.
\end{abstract}

@@

Consider the task of integrating a root growth model with a shoot growth model. The root growth model is written in C (see Listing \ref{lst:src:root}) and can be expressed as
\begin{equation}
R_{t+1} = R_{t} \times r_{r} \times dt + R_{t}.
\end{equation}
The inputs to the root growth model are
\begin{itemize}
	\item $R_{t}$: the root mass at time $t$
	\item $r_{r}$: the relative root growth rate
	\item $dt$: the time step
\end{itemize}
and the output of the root growth model is $R_{t+1}$, the root mass at the next time step. 
%
The shoot growth model is written in Python (see Listing \ref{lst:src:shoot}) and can be expressed as
\begin{equation}
S_{t+1} = S_{t} \times r_{s} \times dt + S_{t} - (R_{t+1} - R_{t}).
\end{equation}
The inputs to the shoot model are 
\begin{itemize}
	\item $S_{t}$: the shoot mass at time $t$
	\item $r_{s}$: the relative shoot growth rate
	\item $dt$: the time step
	\item $R_{t}$: the root mass at time $t$ 
	\item $R_{t+1}$: the root mass at time $t+1$ 
\end{itemize}
and the output of the shoot growth model is $S_{t+1}$, the shoot mass at the next time step.
%
While it is possible to run both models in isolation if all of the appropriate input variables are provided, the value of $R_{t+1}$ calculated by the root growth model could be used as input to the shoot growth model. However, because the two models are written in different languages, the two models cannot be directly integrated. To integrate the models a scientist must either "manually" integrate them by running one model and then the other or translate one of the models into the language of the other so that it can be called directly. For a tightly coupled set of models such as these where one model depends directly on the result from the other at the current time step, manual integration is very inefficient for more than a handful of time steps even when done using a script. Although translating one or both of these simple toy models would be relatively straight forward, actual plant models are much more complex and could contain thousands of lines of code that would be time consuming to translate and result in unnecessary replication of model algorithms.

We present an Open Source Python package, {\pkg}, for creating and running integration networks by connecting existing computational models written in different programming languages. {\pkg} was developed as part of the Crops \emph{in Silico} \citep{Marshall-Colon2017} initiative to build a complete crop \emph{in silico} from the level of the genes to the level of the field. {\pkg} is available on Github\footnote{{\tt https://github.com/cropsinsilico/cis\_interface}} with full documentation\footnote{{\tt https://cropsinsilico.github.io/cis\_interface/}} and can be installed from PyPI\footnote{{\tt https://pypi.org/project/cis\_interface/}} via {\tt pip install {\pkg}} or conda-forge\footnote{{\tt https://github.com/conda-forge/cis\_interface-feedstock}} via {\tt conda -c conda-forge install {\pkg}}. Although developed for the purpose of integrating plant biology models, {\pkg} can be applied to any situation requiring coordination between models in the supported languages. \S\ref{S:background} provides background information on existing efforts for cross-language model integration, \S\ref{S:methods} describes the methods used by the {\pkg} package to integrate models, \S\ref{S:results} presents several tests demonstrating the performance of message passing between integrated models, and \S\ref{S:discuss} summarizes the features of the {\pkg} package, describes a few use cases, and outlines areas of ongoing and future development.

%---------------------------------------------------------------------------------
%---------------------------------------------------------------------------------
% BACKGROUND
\section{Background}\label{S:background}
%
Computational plant models are usually written with a very specific research question in mind \citep[e.g. describing a specific metabolic pathway in C4 photosynthesis,][]{Wang2014}. The resulting model codes are often highly tuned for this one purpose and are written specifically for scientists within an isolated group of collaborators. As a result, it is unlikely that models produced by two independent groups will be directly compatible in terms of language, data format, or units \citep{Marshall-Colon2017}.

There are models written in Matlab \citep{ZHU2013, Wang2014}, Python \citep{Pradal2009}, C++ \citep{Merks2011, Postma2017}, Microsoft Excel \citep{Sharkey2016}, VisualBasic \citep{Humphries1995, HALL2013}, R \citep{WANG2015}, Java \citep{Song2013, Kappas2013}, Fortran \citep{Goudriaan1994}, and several domain specific languages \citep[e.g. SBML,][]{Hucka2003}. The variety of data formats used by these codes is just as diverse. While some specific subfields have settled on standards for things like microarray gene expression data \citep[e.g. MINiML,][]{GEO2017}, many models use unique data formats that may or may not include metadata such as the data types, field names, or units. 
%For example, OpenAlea \citep{Pradal2008} defines its own Multiscale Tree Graph (MTG) tree structure for describing plant topology at different scales that is an extension to the MTG format used by AMAPmod \citep{Godin1998}. 
For example, while LPy \citep{Boudon2012} and Houdini FX \citep{HoudiniFX} both allow users to specify plant structures via L-systems, they differ slightly in their syntax and so are not directly compatible. Or, as another example, 3D canopy structures can be generated in any unit and standard 3D geometry formats like Ply \citep{ply} and Obj \citep{obj} do not include units in their metadata. As a result, if an LPy model produces a canopy in Ply format with units of centimeters, the photosynthetic photon flux density (PPFD) output by a ray tracer like fastTracer \citep{Song2013} (which expects a geometry in units of meters) would be incorrect.

As a result, integrating two biological models like the root and shoot models from \S\ref{S:intro} requires that the following questions be addressed first:
%
\begin{enumerate}
	\item Orchestration: How will models be executed?
	\item Communication: How will information be passed from one model to the next across languages?
	\item Translation: How will data output by one model be translated into a format understood by the next model?
\end{enumerate}

These three questions can be addressed through manual integration (e.g. running the root model, converting the output root mass into the units expected by the shoot model, and then running the shoot model). However, manual integration is 1) computationally inefficient when many iterations are necessary, such as during parameter searches or steady state convergence tests, 2) unique to every integration, requiring the production of new scripts, and 3) complex when more than three models are being integrated. Within both computational biology and the larger scientific computing community, groups have developed around different solutions to the problem of connecting models. These solutions can generally be grouped into two categories: domain specific languages and frameworks.

@@

{\pkg} is an example of a framework that overcomes these issues by exposing simple and easily accessible interfaces in the language of the model (See \S\ref{SSS:interface}) that permit messages to be passed between the model processes as they run in parallel (See \S\ref{SS:communication}). As a result, models writers only need knowledge of the language in which their model is written.

@@

{\pkg} was designed to address the questions from \S\ref{S:background} while being:

@@

{\pkg} currently support models written in Python, Matlab, C, and C++ with 
additional Domain Specific Language (DSL) support for LPy models \citep{Boudon2012}. Support for additional languages is planned for future development (See \S\ref{SS:langsupport}).

%---------------------------------------------------------------------------------
% ORCHESTRATION
\subsection{Orchestration}\label{SS:orchestration}
%
{\pkg} is executed via a command line interface (CLI), {\pkgrun} with specification files (\S\ref{SSS:yaml}) as input. Based on the information contained in the specification files, {\pkg} establishes a network of asynchronous communication channels (\S\ref{SS:communication}), and launches the models on new processes (\S\ref{SS:model_execution}).
%At execution, {\pkg} launches models on new processes based on information contained in a YAML specification file (See \S\ref{SSS:yaml} below) and establishes

@@

Model YAMLs include information about the location of the 
model source code, the language the model is written in, how the model should 
be run, and any input or output variables including their data type and units. 
%
%For example, Listings \ref{lst:yml:root} and \ref{lst:yml:shoot} from \A\ref{A:yml} are the model YAMLs for the example root and shoot growth models from \S\ref{S:intro}. These model YAMLs could be used in any integration without being modified because the model inputs and outputs will not change.
%
%\rootyml
%
%\shootyml
%
Integration networks are specified by declaring the connections 
between models and connections are declared by pairing an output variable from 
one model with the input variable of another model. The {\pkg} 
CLI sets up the necessary communication mechanisms to 
then direct data from one model to the next in the specified pattern. Models 
can have as many input and/or output variables as is desired and connections between models are specified by references to the 
input/output variables associated with each model. In addition, input and output variables can also be connected to files.
%
%For example, in the connection YAML for integrating the root and shoot models (Listing \ref{lst:yml:int} from \A\ref{A:yml}), the first two connections link the {\tt root\_input} and {\tt shoot\_input} input channels from the root and shoot model YAML's to input files. The third connection links the {\tt next\_root\_mass} output variable from the root model YAML (Listing \ref{lst:yml:root}) to the {\tt next\_root\_mass} input variable from the shoot model YAML (Listing \ref{lst:yml:shoot}). The forth connection links the {\tt next\_shoot\_mass} to an output file.
%
%\intyml
%
This format allows users the 
flexibility to create complex integration networks and test models in isolation before running the entire integration network.

%---------------------------------------------------------------------------------
% MODEL EXECUTION
\subsubsection{Model Execution}\label{SS:model_execution}
%
{\pkg} launches each model in an integration on its own process, allowing models to complete independent operations in parallel and complete tasks more quickly. For example, the root and shoot growth models from \S\ref{S:intro} require 2.42\,s and 4.06\,s respectively to run for 100 time steps in isolation. If these two models were manually integrated in serial via the command line, the integration would require a total of 6.48\,s. However, the same integration requires only 4.52\,s using {\pkg}, a speed-up of 1.43. The exact speedup provided by parallel integration using {\pkg} depends on how independent the models are and how equally distributed the work is between the models. If the models are fully independent, the speedup is limited by the time required to execute the slowest model. In the example, this would be the shoot growth model which required almost twice the time to execute as the root model. If the models are dependent on one-another, the models will have to wait for messages to be passed and the speedup will not be as great.

@@

For the compiled languages (C and C++) there are a few options. The user can 
compile the model themselves, provided they link against the appropriate 
{\pkg} header library. Alternatively, users can provide the location of 
the model source code and let {\pkg} handle the compilation, including 
linking against the appropriate {\pkg} interface library. {\pkg} 
also has support for compiling models using Make \citep{Stallman2004} and 
CMake \citep{Martin2006} for 
models that already have a Makefile or CMakeLists.txt. To use these tools to compile a model, 
lines are added to the recipe in order to allow linking against {\pkg}.

Once the models are running, {\pkg} uses threads on the master process to monitor the progress of the model and report back model output (e.g. log messages printed to stdout or stderr) and any status changes. If a model issues any errors, the master process will shut down any model processes that are still running and close any connections, discarding any unprocessed messages. If a model completes without any errors, the master process will cleanup any connections that are no longer required after waiting for all messages to be processed. The master process will only complete once an error is encountered or all model processes have completed.

%---------------------------------------------------------------------------------
% COMMUNICATION
\subsection{Communication}\label{SS:communication}
%
In order for integrations to be possible, models must be able to send and receive information to and from other models written in different languages via some communication mechanism. 
Communication within {\pkg} integration networks was designed to be flexible in terms of the 
languages, platforms, and data types available. To accomplish this, {\pkg} leverages 
three different tools for communication, 
System V IPC Queues (\S\ref{SSS:ipc}), ZeroMQ (\S\ref{SSS:zmq}), and RabbitMQ (\S\ref{SSS:rmq}), 
which each have their own strengths and weaknesses. The 
particular communication mechanism used by {\pkg} is determined by the platform, available libraries, and integration strategy. However, regardless of the communication mechanisms, the user will always use the same interface in the language of their model (See \S\ref{SSS:interface}).

%---------------------------------------------------------------------------------
% SYSTEM V IPC
\subsubsection{System V IPC Queues}\label{SSS:ipc}
%
The first communication mechanism used by {\pkg} was System V interprocess 
communication (IPC) message queues \citep{Rusling1999} on Posix (Linux and Mac OS X) systems. 
IPC message queues allow messages to be passed between models running on separate 
processes on the same machine. While IPC message queues are simple, fast 
(See \S\ref{SS:results_commtype}), and are built into most Posix operating systems, they do not work in all 
situations. IPC queues are not natively supported by Windows operating systems 
and do not allow communication between remote processes. In addition, IPC queues also 
have relatively low default message size limits on Mac OS X systems (2048 bytes). Once 
the queue is full or if the message is larger than the limit, any process attempting to send an additional message will block until a 
sufficient number of messages has been removed from the queue to accommodate the new message. 
For messages larger than the limit, the sending process will block indefinitely. 
This can be handled by splitting large messages into multiple smaller messages 
(See \S\ref{SSS:large}), however, the time required to send a message increases with the number of 
message it must be broken into (See \S\ref{SS:results_commtype}). These limits make sending large messages 
relatively inefficient when compared with other communications mechanisms. 
As a result, IPC queues are used by 
{\pkg} as a fallback on Posix systems if none of the other supported communication libraries have 
not been installed.

%---------------------------------------------------------------------------------
% ZEROMQ
\subsubsection{ZeroMQ}\label{SSS:zmq}
%
The preferred communication mechanism used by {\pkg} are ZeroMQ sockets 
\citep[ZMQ][]{Akgul2013}. ZMQ 
provides broker-less communication via a number of protocols and patterns with 
bindings in a wide variety of languages that can be installed on Posix and Windows 
operating systems. ZMQ was adopted by {\pkg} in order to allow support 
on Windows and for future target languages (See \S\ref{SS:future}) that could not be 
accomplished using IPC queues. In addition, while ZMQ allows 
interprocess communication via IPC, ZMQ also 
supports protocols for distributed communication via an Internet Protocol (IP) 
network. While {\pkg} does not currently support using these protocols 
for distributed integration networks, this is a future development plan that has been prepared for.

%---------------------------------------------------------------------------------
% RABBITMQ
\subsubsection{RabbitMQ}\label{SSS:rmq}
%
While broker-less communication like ZMQ is light weight and fast, it is not as fault 
tolerant as brokered messaging systems that confirm message delivery and can resend 
dropped messages. Resilience to dropped messages, while not as necessary for integrations 
running entirely on a local machine, will be more important for integrations running on 
distributed resources with less reliable connections. As a result, {\pkg} includes support 
for brokered communication via RabbitMQ \citep[RMQ][]{RMQ} that will be used during future 
development to allow integrations to run on distributed resources or include remote models 
run as services (\S\ref{SS:distributed}). Due to the slower message speed, 
{\pkg} does not currently use RMQ for communication unless explicitly specified by the user in their 
integration network YAML.

%---------------------------------------------------------------------------------
% ASYNCHRONOUS MESSAGE MASSING
\subsubsection{Asynchronous Message Passing}\label{SSS:async}
%
The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\pkg}. 
Models do not block on sending messages to 
output channels; the model is free to continue working on its task while 
an \emph{output driver} waits for the message to be routed and received on a separate master process thread. Similarly, 
an \emph{input driver} continuously pings input channels on another thread, moving received messages into a 
buffer queue so that they are ready and waiting for the receiving model 
when it asks for input. These drivers work together to move message along from one model to the next like a conveyor belt. 
As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved. Figure \ref{fig:async} describes the general flow of messages.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/io_drivers.png}
	\caption{Diagram of how messages are passed asynchronously using input/output drivers and an intermediate channel.}
	\label{fig:async}
	\end{center}
	\end{figure}
\fi
%
\begin{enumerate}
	\item A model sends a message in the form of a native data object via a language-specific API to one of the output channels declared in the model YAML. The output channel interface encodes the message and sends it.
	\item An output connection driver (written in Python) runs in a separate thread on the master process, listening to the model output channel. When the model sends a message, the output connection driver checks that the message is in the expected format and then forwards it to an intermediate channel. The intermediate channel is used as a buffer for support on distributed architectures (e.g. if one model is running on a remote machine). In these cases, the intermediate channel will connect to a RMQ broker using security credentials.
	\item An input connection driver (also written in Python) runs in a separate thread, listening to the intermediate channel. When a message is received, it is then forwarded to the input channel of the receiving model as specified in the integration network YAML.
	\item The receiving model receives the message in the form of an analogous native data object. Interface receive calls can be either blocking or non-blocking, but are blocking by default.
\end{enumerate}
